---
title: "Biomath HW02"
author: "Maxwell Greene"
date: "February 13, 2020"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 6,fig.height=4)
library(mosaic)
library(mosaicCalc)
library(deSolve)
library(ggplot2)
```

#Problem #1
Consider the equation
$$
\frac{dN}{dt}= \kappa - rN \text{,  where } \kappa , r>0
$$

##a)
If the populaton is above $\frac{\kappa}{r}$, it will decrease until it reaches $\frac{\kappa}{r}$. If it below it will increase until it reaches $\frac{\kappa}{r}$. Notably, the maximum increase of the function is $\kappa$, whereas there is no limit to the rate of decrease.  
Equivalently,
$$
\frac{dN}{dt}>0 \text{ if } \kappa-rN>0, \frac{\kappa}{r}>0 \\
\frac{dN}{dt}<0 \text{ if } \kappa-rN<0, \frac{\kappa}{r}<0
$$
This implies a constant birth rate, $\kappa$, and a death rate proportional to the populaton, scaled by the factor $r$.


##b)
The population will stabilize at $N=\frac{\kappa}{r}$ for the reasons in (a).

##c)
$$
\begin{aligned}
\frac{dn}{dt} &= \kappa -rN \\
\int \frac{1}{\kappa -rN}dN &= \int dt \\
-\frac{1}{r} \ln(\kappa -rN)&=t+c \\
N(t)=-\frac{1}{r} \Big( e^{r(t+c)}-\kappa \Big) &
\text{, } \quad N(0)=N_0 \\
\therefore \quad c = -\frac{1}{r}\ln \Big( \kappa -rN\Big)
\end{aligned}
$$

##d)
Yes, this solution does fit the results calculated above.
See below for justification
```{r}
r <- .75; k <- 10; N0 <- 10; c <- -(1/r)*(log(k-r*N0));
t <- seq(0,10,length.out=1001)
y <- -(1/r)*(exp(-r*(t+c))-k)
plot(t,y)
```

#Problem #2
$$
\begin{aligned}
X(t)&=X_\infty+(X_0-X_\infty)e^{-\lambda t} \\
X(0)&=X_\infty+(X_0-X_\infty)e^0 = X_\infty+X_0-X_\infty = X_0 \\
\lim_{t \rightarrow \infty} X(t) &= X_\infty -\lambda (X_0-X_\infty)e^{-\infty} = X_\infty \\
\frac{dX}{dt} &= X_\infty - \lambda(X_0-X_\infty)e^{-\lambda t}
\end{aligned}
$$

#Problem #3
```{r}
func <- dC ~ qi*(v-C)/(v0+(qi-q0)*t)
solN <- integrateODE(func,v0=100,v=35,qi=5,q0=1,C=5,tdur=list(from=0,to=10))
plotFun(solN$C(t)~t,ylim=range(4,16),t.lim=range(0,10))
```

#Problem #4
System of equations for figure #1:
$$
\begin{aligned}
\frac{dX_1}{dt}&=L-x_1(a_{01}+a_{21}+a_{31})+x_3a_{13}+x_2a_{12}-x_1a_{01}\\
\frac{dX_2}{dt}&=x_1a_{21}-x_2(a_{02}+a_{12}) \\
\frac{dX_3}{dt}&=x_1a_{31}-x_3a_{13}
\end{aligned}
$$

#Problem #5

#Problem #6























