---
title: "BiomathHW07"
author: "Maxwell Greene"
date: "April 17, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=5,fig.height = 3.2)
library(phaseR)
library(deSolve)
library(mosaic)
```

#Problem #1
The following is a generic code to classify the behavior of a linear system based on it's trace and eigenvalues. I will use it in each part of the problem 

```{r}
classify <- function(mat)
{
  trace <- sum(diag(mat)); deter <- det(mat)
  delta <- deter - trace^2/4
  if(deter < 0){return("saddle")}
  if(deter ==0)
  {
    if(trace > 0){return("unstable line")}
    if(trace < 0){return("stable line")}
    if(trace ==0){return("uniform motion")}
  }
  if(trace > 0)
  {
    if(delta < 0){return("source")}
    if(delta ==0){return("degenerate source")}
    if(delta > 0){return("spiral source")}
  }
  if(trace ==0){return("center")}
  if(trace < 0)
  {
    if(delta < 0){return("sink")}
    if(delta ==0){return("degenerate sink")}
    if(delta > 0){return("spiral sink")}
  }
}
plotSystem <- function(mat)
{
  simple <- function(t,state,parameters){
    with(as.list(c(state,parameters)),{
    dx <- a*state[1] + b*state[2]
    dy <- c*state[1] + d*state[2]
    list(c(dx,dy))
  })}
  ff  <- flowField(simple,
                   xlim = c(-2, 2), ylim = c(-2, 2),
                   parameters = c(a=mat[1],b=mat[3],c=mat[2],d=mat[4]),
                   points = 19,add = FALSE)
  state <- matrix(c(1,1,1,-1,-1,1,-1,-1,0,2,0,-2,-1,0,1,0),
                                    8, 2, byrow = TRUE)
  trajs <- trajectory(simple,y0 = state, tlim = c(0, 10),
                      parameters = c(a=mat[1],b=mat[3],c=mat[2],d=mat[4]),add=TRUE)
}
```

##(a)
$$
\begin{aligned}
\dot{x}&=6x+2y \\
\dot{y}&=2x+3y
\end{aligned}
$$

```{r,echo=FALSE,message=FALSE}
A <- matrix(c(6,2,2,3),nrow = 2)
print(paste("Phase portrait type: ",classify(A)))
```
Therefore we expect to have two distinct real eigenvalues:
```{r,echo=FALSE,message=FALSE}
print(paste("Eigenvalues: ",eigen(A)$value))
plotSystem(A)
```

##(b)
$$
\begin{aligned}
\dot{x}&=x+2y \\
\dot{y}&=4x+3y
\end{aligned}
$$

```{r,echo=FALSE,message=FALSE}
B <- matrix(c(1,4,2,3),nrow = 2)
print(paste("Phase portrait type: ",classify(B)))
```
Therefore we expect two distinct real eigenvalues:
```{r,echo=FALSE,message=FALSE}
print(paste("Eigenvalues: ",eigen(B)$value))
plotSystem(B)
```

##(c)
$$
\begin{aligned}
\dot{x}&=-2x+4y \\
\dot{y}&=-x+y
\end{aligned}
$$

```{r,echo=FALSE,message=FALSE}
C <- matrix(c(-2,-1,4,1),nrow = 2)
print(paste("Phase portrait type: ",classify(C)))
```
Therefore we expect a complex conjugate pair of eigenvalues:
```{r,echo=FALSE,message=FALSE}
print(paste("Eigenvalues: ",eigen(C)$value))
plotSystem(C)
```

##(d)
$$
\begin{aligned}
\dot{x}&=2x+y \\
\dot{y}&=x+2y
\end{aligned}
$$

```{r,echo=FALSE,message=FALSE}
D <- matrix(c(2,1,1,2),nrow = 2)
print(paste("Phase portrait type: ",classify(D)))
```
Therefore we expect two distinct real eigenvalues:
```{r,echo=FALSE,message=FALSE}
print(paste("Eigenvalues: ",eigen(D)$value))
plotSystem(D)
```

##(e)
$$
\begin{aligned}
\dot{x}&=x+5y \\
\dot{y}&=x-3y
\end{aligned}
$$

```{r,echo=FALSE,message=FALSE}
E <- matrix(c(1,1,5,-3),nrow = 2)
print(paste("Phase portrait type: ",classify(E)))
```
Therefore we expect two distinct real eigenvalues:
```{r,echo=FALSE,message=FALSE}
print(paste("Eigenvalues: ",eigen(E)$value))
plotSystem(E)
```

##(f)
$$
\begin{aligned}
\dot{x}&=-1x+ay \\
\dot{y}&=0x+ay \\
&\text{for }a \neq 0
\end{aligned}
$$
This function changes behavior, dependent on the value of $a$, when it crosses $det(f)=0, tr(f)=0, det(f)=\frac{tr(f)^2}{4}$ lines on the trace-determinant plane. So I will find these critical values alegbraically.
$$
\begin{aligned}
tr(f)&=-1+a=0 \quad \rightarrow \quad \boxed{a=1} \\
&a=1 \text{ is not critical value since } det(d)<0\\
det(f)&= -a=0 \quad \rightarrow \quad \boxed{a=0} \\
delta&=det(f)-\frac{tr(f)^2}{4}=-a-\frac{(a-1)^2}{4} \\
&=a^2+2a+1 \quad \rightarrow \quad \boxed{a=-1}
\end{aligned}
$$

Classification and visualization for $a=-1.5,-1.0,-0.5,0,0.5$:
```{r,collapse=TRUE}
f_n1.5 <- matrix(c(-1,0,-1.5,-1.5),nrow = 2)
f_n1.0 <- matrix(c(-1,0,-1.0,-1.0),nrow = 2)
f_n0.5 <- matrix(c(-1,0,-0.5,-0.5),nrow = 2)
f_p0.0 <- matrix(c(-1,0,+0.0,+0.0),nrow = 2)
f_p0.5 <- matrix(c(-1,0,+0.5,+0.5),nrow = 2)
```
```{r}
classify(f_n1.5)
classify(f_n1.0)
classify(f_n0.5)
classify(f_p0.0)
classify(f_p0.5)
```
Therefore we expect the following eigenvalue combinations (in order):  
-Distinct real  
-One real  
-Distinct real  
-Distinct real  
-Distinct real  
```{r,echo=FALSE}
print(paste("Eigenvalues: ",eigen(f_n1.5)$value))
print(paste("Eigenvalues: ",eigen(f_n1.0)$value))
print(paste("Eigenvalues: ",eigen(f_n0.5)$value))
print(paste("Eigenvalues: ",eigen(f_p0.0)$value))
print(paste("Eigenvalues: ",eigen(f_p0.5)$value))
```

```{r}
fs <- list(f_n1.5,f_n1.0,f_n0.5,f_p0.0,f_p0.5)
deters <- lapply(fs,det)
traces <- lapply(fs,function(A){sum(diag(A))})
vals <- seq(from = -2.5,to=2.5,length.out=101)
plot(vals,vals^2/4,col="blue",
     xlim=c(-2.5,-.5),ylim=c(-0.5,1.5),
     type="l",xlab="trace",ylab="determinant")
points(vals,rep(0,101),col="blue",type="l")
points(traces,deters,col="red")
```

Therefore, the system exhibits the following behavior:  
Sink for $(- \infty<a<-1) \cup (-1,0)$  
Degenerate sink for $a=-1$  
Stable line for $a=0$  
Saddle for $(0<a<\infty)$  





